### 2020-12-18 msc-thesis.tex
Introduction

Many tools require identification of a user in case interactions or initiated business processes have to be associated with him. An identity provider (IdP) who stores partial identities of users therefore is part of numerous system architectures. This leads to customers of SPs to be associated with multiple partial identities - one for each SP. Service Providers usually enable users to manage their partial identities using online self-service through so called User Profiles.

The issues of this approach becomes clear for a large amount of partial identities. The more Service Providers a person is customer of, the more partial identities he will have to manage through user profiles. Changing the home address in each profile becomes a major inconvenience. More serious problems occur for example through the increased risk of data breaches.

An improved model for identity management is necessary. In recent years, multiple new models have been invented. One model called Social Login became popular. It enables customers to leverage existing User Profiles of social networks like Facebook to create new User Profiles for Service Providers. This improves identity management by enabling SPs to rely on the identity information they are provided by the external profile. Users can manage identity related information through their Social Profile and each change would then be propagated to User Profiles of connected SPs.

This approach, however, still leaves the requirement of User Profiles for each SP because depending on the provided service, additional attributes which are not part of the Social Profile, have to be manageable by the user. This includes for example subscriptions or contracts.

The solution to this problem is an Identity Management System (IMS) which manages the whole online identity of a person and can replace the individual partial identities and User Profiles. The system itself is provided and hosted by a SP for consumption by other SPs. To replace existing User Profiles, the IMS can not only rely on systems for registration, authentication, authorisation and provisioning. Additional requirements of SPs like communication, data wallet and management of additional attributes have to be satisfied.

In order for an SP to use identities and services of an IMS, an integration into the system architecture of the SP has to take place. Due to the complex nature of system architectures and identity management, this is a difficult task.

A currently relevant example for usage of online self-service with particular interesting requirements for identities is the "Online Access Law" (OZG). It requires all administrative services of the German federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make the profiles only available for usage in context of the OZG. From a user perspective, this would be yet another partial identity and user profile to manage.

\chapter{Objective}
Based on the prominent example of the OZG, the bachelor thesis will construct a message based integration architecture which enables system architectures of service providers with existing user profiles to make their services usable with an Identity Management System. In order to make the integration architecture suitable to real life requirements, the OZG is selected as an example due to its current relevance and high requirements regarding identity management. The integration architecture is however not only applicable in cases relevant for OZG but also for other Service Providers with similar requirements. As requirements of SPs are not guaranteed to be similar to those of the OZG, the integration architecture is built to be expendable.

The \textbf{integration architecture} for the OZG enables the \textbf{IMS functionalities} to be usable for selected \textbf{OZG scenarios} by integrating a \textbf{connector} in the \textbf{system architecture} of a member state.

\begin{itemize}
    \item OZG scenarios are based on documented digitalization efforts of administrative services in the context of the OZG and are for example be the application for an administrative service or the communication with a governmental institution.
    \item The system architecture is based on documentation about architecture plans of member states.
    \item The functionalities of the IMS are based on a requirements analysis of OZG scenarios and additional research.
    \item The connector is defined based on functionalities the integration architecture requires from the IMS.
    \item The integration architecture is the main scientific contribution of this bachelor thesis. It is based on messaging patterns in order to utilize a combination of established and tested messaging solutions. In order to save investments it also reuses as many system components as possible and modifies as few system components as necessary.
\end{itemize}

\chapter{Structure of Work}
In the beginning of the thesis, basic concepts and terminology is explained on which the following chapters will rely.

Then, preparatory work relevant for the integration architecture is presented.
Based on process diagrams of digitalized administration services, several use cases of OSS are listed and evaluated for their relevance. They provide important information on which functionalities the IMS is required to have and the integration architecture is required to integrate.

An integration approach through a connector is used. Concrete functionalities and interfaces of a possible connector are specified for usage by the integration architecture.

Based on documentation of system architectures of several member states, relevant components of the architecture relevant for the OZG are selected. Each component is explained in detail along with the functionalities it provides for usage by the integration architecture.

The integration architecture integrates previously documented connector and components of the system architecture in order to make the selected OSS cases usable with the IMS. The architecture is presented iteratively. Each iteration contains increasingly demanding requirements along with the documentation of the increasingly complex integration architecture. Each documentation is done through textual explanations, multiple flow diagrams and one messaging architecture.

In order to validate the integration architecture, the execution of one administration service with usage of the IMS is evaluated in theory.

@@

This chapter defines many of the terms which were used in the previous introduction and further elaborates the underlying concepts and relevance for the objective of the bachelor thesis.

\section{Service Provider}
Service providers (SP) are entities like enterprises or organisations which make services accessible over the internet.

\section{Online Self-Service}
Online self-service (OSS) describes a method used by Service Providers to make their services available to the user. The method is characterized by being available over the internet, requiring no direct human interaction but instead enabling the user to access services in a self-reliant way. The user is usually supported by a variety of software solutions.

\section{Self-Service Tools}
Self-service tools are software solutions for customers to enable a online self-service experience. Depending on the use case, the tools can be transactional or informing. Informing tools help users to retrieve relevant information, often replacing human customer support. Transactional tools help users to interact with services in a persistent way to for example save personal data or trigger business processes.

@@

\section{Identity}

\section{User Profile}

\section{Identity Management System}

\section{OZG}

This section introduces the law for improvement of online access of administrative services (Online Access Law - OZG), its origin, purpose and planned realization.

\subsection{Context}

The Federal Ministry of the Interior, Building and Community is, amongst other things, responsible for modernizing public administration \cite{BMI:Moderne_Verwaltung}. Current modernization efforts lie in the construction of an E-Government \cite{BMI:Behoerdengaenge}.
The E-Government-Law in 2013 was an early step increasing internal usage of digital systems in governmental administration. This includes for example the usage of digital personal files and scanning of documents for replacement \cite{BMI:E-Government_Gesetz}. 
In 2017 the German government passed the law for improvement of online access to administrative services (OZG) which focuses on usage of administration services by the people \cite{BMI:Onlinezugangsgesetz}. 
Modernization of the administration, however, is not exclusive to Germany as in 2018 the European Parliament and Council decided on a Single Digital Gateway (SDG) providing uniform access to digital administrative governmental services of every European country \cite{BMI:Single_Digital_Gateway}.

\subsection{Motivation}

@@

\subsection{The OZG}

@@

\subsection{OZG Execution}

@@

Many services, which are still bound to paper must be digitalized. In total, the BMI lists 575 relevant services. Some of them are provided by the federal republic, some by the member states and yet other by the communes. Respective to these responsibilities, the task of digitalization is distributed between federal republic and member states, where each member state is assigned to take the lead for specific areas. \cite{BMI:Onlinezugangsgesetz} Results of the digitalization are documented on the "OZG-Informationsplattform" \cite{BMI:Informatiosplattform} as process diagrams and data schemata.

@@

@@

\section{OZG OSS Scenarios}

The information platform of the OZG contains information about many administration services in form of process diagrams \cite{BMI:Ergebnisse}. They serve as a resource to identify important OSS scenarios of administration processes. The following scenarios were identified:

@@

\section{OSS Connector}

\section{OZG System Architecture}
This section describes the system architecture of a member state relevant for the execution of the OZG an integration of a OSS provider. Due to a lack of detailed information, the system architecture is described by components responsible for a category of tasks. Each component is associated to data objects it processes. The definition of components and data objects are based on research about planned system architectures of member states.

@@

@@

This chapter presents a message based integration architecture which enables access to OSS features of member states through a OSS provider. Multiple versions of an integration architecture are presented which fulfill increasingly complex requirements. Requirements are formulated as user scenarios which should be accomplishable through the OSS provider.
Each version is presented in the form of multiple flow charts, each describing sequential steps while carrying out a required scenario and a corresponding messaging infrastructure.

>>>>

Many tools require identification of a user in case interactions or initiated business processes have to be associated with him. A system for management of identities of users therefore is part of numerous system architectures. Service Providers usually enable users to manage their partial identities using online self-service through so called User Profiles. Today, customers of SPs are associated with multiple partial identities and user profiles - one for each SP

The result is a digital identity crisis \cite{IdentityCrisis}:
\begin{itemize}
    \item Multiple SPs collect, use and share parts of identities which leads to a distributed and fragmented identity.
    \item As a result of the distributed identity, risk of data breaches increases.
    \item Managing multiple use profiles is very inconvenient for users.
    \item It is often unclear who stores which data and for what purpose.
    \item SPs have an incentive to collect and share more user data than they need, to provide their service.
\end{itemize}

An improved model for identity management is necessary. In recent years, multiple new models have been invented. One model called Social Login became popular. It enables customers to create User Profiles for SPs through existing User Profile of Social Networks like Facebook. This improves identity management by enabling SPs to rely on the external profile for unique identification, authorisation and personal attributes. Users can therefore manage identity related information through their Social Profile which is accessed by the SP.

This approach, however, still leaves the requirement of User Profiles for each SP because, depending on the provided service, additional attributes which are not part of the Social Profile, have to be manageable by the user. SPs also might require additional online self-service tools the social network does not provide.

The solution to this problem is an Identity Management Provider (IMP) which manages the whole online identity of a person and can replace the individual partial identities and User Profiles. The system itself is provided and hosted by a SP for consumption by other SPs. To replace existing User Profiles, the IMP can not only rely on systems for registration, authentication, authorisation and provisioning. Additional requirements of SPs like communication, data wallet and management of additional attributes have to be satisfied.

In order for an SP to use identities and services of an IMP, an integration into the system architecture of the SP has to take place. Due to the complex nature of system architectures and identity management, this is a difficult task.

A currently relevant example for usage of online self-service with particular interesting requirements for identities is the German "Online Access Law" (Online Zugangs Gesetz: OZG). It requires all administrative services of the German federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make the profiles only available for usage in context of the OZG. From a user perspective, this would be yet another partial identity and user profile to manage.

\chapter{Objective}
Based on the prominent example of the OZG, the bachelor thesis will construct a message based integration architecture which enables system architectures of service providers with existing user profiles to make their services usable with an Identity Management Provider. In order to make the integration architecture suitable to real life requirements, the OZG is selected as an example due to its current relevance and high requirements regarding identity management. The integration architecture is however not only applicable in cases relevant for OZG but also for other Service Providers with similar requirements. As requirements of SPs are not guaranteed to be similar to those of the OZG, the integration architecture is built to be expendable.

The \textbf{integration architecture} for the OZG enables \textbf{IMP functionalities} to be usable for selected \textbf{OZG scenarios} by integrating a \textbf{connector} in the \textbf{system architecture} of a member state.

\begin{itemize}
    \item OZG scenarios are based on documented digitalization efforts of administrative services in the context of the OZG and are for example be the application for an administrative service or the communication with a governmental institution.
    \item The system architecture is based on documentation about architecture plans of member states.
    \item The functionalities of the IMP are based on a requirements analysis of OZG scenarios and additional research.
    \item The connector is defined based on functionalities the integration architecture requires from the IMP.
    \item The integration architecture is the main scientific contribution of this bachelor thesis. It is based on messaging patterns in order to utilize a combination of established and tested messaging solutions. In order to save investments it also reuses as many system components as possible and modifies as few system components as necessary.
\end{itemize}

\chapter{Structure of Work}
In the beginning of the thesis, terminology and basic concepts are explained.
This includes definitions of terms like Online Self-Service, Identity and Identity Management. The concept of Identity Management Providers and the Online Access Law are explained. The special relevance for integration of IMPs and OZG are described.

Then, preparatory work relevant for the integration architecture is presented. Relevant OZG scenarios based on digitisation documentations of administrative services are analyzed in order to define important features the IMP is required to have. The connector which makes the IMP accessible towards the integration architecture is documented. Based on documentation of system architectures of several member states, relevant components of the architecture relevant for the OZG are selected. Each component is explained in detail along with the functionalities it provides for usage by the integration architecture.

The integration architecture integrates previously documented connector and components of the system architecture in order to make selected OZG scenarios usable with the IMP. The architecture is presented iterative. Each iteration contains increasingly demanding requirements along with the documentation of the increasingly complex integration architecture. Each documentation is done through textual explanations, multiple flow diagrams and one messaging architecture.

In order to validate the integration architecture, the execution of one administration service with usage of the IMP is evaluated in theory.

@@

\section{Terminology}

\subsection{Service Provider}
Service providers (SP) are entities like enterprises or organisations which make services accessible over the internet.

\subsection{Online Self-Service}
Online self-service (OSS) describes a method used by Service Providers to make their services available to the user. The method is characterized by being available over the internet, requiring no direct human interaction but instead enabling the user to access services in a self-reliant way. The user is usually supported by a variety of software solutions.

\subsection{Self-Service Tools}
Self-service tools are software solutions for customers to enable a online self-service experience. Depending on the use case, the tools can be transactional or informing. Informing tools help users to retrieve relevant information, often replacing human customer support. Transactional tools help users to interact with services in a persistent way to for example save personal data or trigger business processes.

@@

\subsection{Digital Identity}
The digital identity, for simplicity called identity in the thesis, is the sum of all digital personal information. This includes a set of identifiers which are pieces of information that can uniquely identify an entity and attributes which describe properties of an entity but do not uniquely identify it. It is also possible to use an aggregation of attributes as identifier.

Identifiers which are commonly used by SPs are the phone number and E-Mail address. Attributes are for example nicknames, hobbies, interests or education. The full name of a person is often not sufficient unique identification, the aggregation of name, age and home address can therefore be used as identifier.

\subsection{Partial Digital Identity}
A partial digital identity is a subset of an identity. Partial identities are commonly used by SPs to manage identity information which is relevant for them.

\subsection{Identity Management}
Identity management is the creation, utilization / reading , updating and deletion of identities or partial identities. The possible utilization of identities depends on the system for management of identities. It could be for example authorization for a service or communication through an inbox.

\subsection{Identity Management System}
Identity management systems are solutions which enable SPs and customers to manage identities.

\subsection{User Profile}
User profiles are one method for identity management. They are characterized by each SP providing a separate identity management system which often consist of an identity provider (IdP) for storing identities, a customer relation management system (CRM) for accessing identities internally and a website with online self-service tools for providing access to customers.

\section{Digital Identity Crisis}
The current situation of digital identities has the following core issues:
\begin{itemize}
    \item aktuelle übersicht über identitiy management
    \begin{itemize}
        \begin{itemize}
            \item oAuth
            \item saml
        \end{itemize}
        \item Social Login
        \item central idps
        \item idendity federation (SOcial login, Single sign on)
        \item device identity kopplung
        \item SSI
    \end{itemize}
\end{itemize}

\section{Identity Management Provider}
Identity management providers (IMP) are a method for identity management which aims to solve the digital identity crisis. They are characterized by giving SPs access to customer identities for which they provide one-stop online self-service. The important aspect of IMP is, that the identity is stored and managed in one place and its utilization is enabled through integration with system architectures of multiple SPs.

\subsection{Objective}
The IMP method can solve the identity issues the following way:

\subsection{IMP System}
A system for IMP can work as follows:

Additional requirements for IMP systems can exist and are analyzed as part of the preparatory work.

\subsection{Integration}
The IMP system has to be integrated. A connector is used:

\subsection{Challenges}
The IMP system has the following problems / open questions regarding integration:
\begin{itemize}
    \item which personal information resides at SP systems an which at IMP systems? => data the user personally manages resides at IMP systems, derived data etc. at SP systems
    \item should data be mirrored at the SP and just a synchronisation functionality exist
\end{itemize}

\section{Online Access Law}
The Federal Ministry of the Interior, Building and Community is, amongst other things, responsible for modernizing public administration \cite{BMI:Moderne_Verwaltung}. Current modernization efforts lie in the construction of an E-Government \cite{BMI:Behoerdengaenge}.
The E-Government-Law in 2013 was an early step increasing internal usage of digital systems in governmental administration. This includes for example the usage of digital personal files and scanning of documents for replacement \cite{BMI:E-Government_Gesetz}. 
In 2017 the German government passed the law for improvement of online access to administrative services (OZG) which focuses on usage of administration services by the people \cite{BMI:Onlinezugangsgesetz}. 
Modernization of the administration, however, is not exclusive to Germany as in 2018 the European Parliament and Council decided on a Single Digital Gateway (SDG) providing uniform access to digital administrative governmental services of every European country \cite{BMI:Single_Digital_Gateway}.

@@

@@

@@

Many services, which are still bound to paper must be digitized. In total, the BMI lists 575 relevant services. Some of them are provided by the federal republic, some by the member states and yet other by the communes. Respective to these responsibilities, the task of digitalization is distributed between federal republic and member states, where each member state is assigned to take the lead for specific areas. \cite{BMI:Onlinezugangsgesetz} Results of the digitalization are documented on the "OZG-Informationsplattform" \cite{BMI:Informatiosplattform} as process diagrams and data schemata.

@@

The OZG is a fitting example for integration of IMP solutions and could benefit from it because:

@@

\section{OZG Processes}

\begin{itemize}
    \item selecting important processes (priority 1)
    \item describing processes: purpose, result, etc
    \item process diagram visualizing
    \item 
\end{itemize}

\section{OZG Scenarios}

\begin{itemize}
    \item structuring parts of upper processes into clusters
    \item 
\end{itemize}

The information platform of the OZG contains information about many administration services in form of process diagrams \cite{BMI:Ergebnisse}. They serve as a resource to identify important OZG scenarios of administration processes. The following scenarios were identified:

@@

\section{OZG System Architecture}
This section describes the system architecture of a member state relevant for the execution of the OZG and integration of an IMP. Due to a lack of detailed information, the system architecture is described by components responsible for a category of tasks. Each component is associated to data objects it processes. The definition of components and data objects are based on research about planned system architectures of member states.

@@

\section{IMP System}
List of functionalities of the IMP System based on research and analysis of the OZG scenarios:

\section{IMP Connector}
Interfaces of the connector for usage by the integration architecture:

\section{Integration Thoughts}

\subsection{OZG}

\subsection{IMP}

@@

This chapter presents a message based integration architecture which enables access to OZG features of member states through an IMP. Multiple versions of an integration architecture are presented which fulfill increasingly complex requirements. Requirements are formulated as user scenarios which should be achievable through the IMP.
Each version is presented in the form of multiple flow charts, each describing sequential steps while carrying out a required scenario and a corresponding messaging infrastructure.
